---
title: 'Data Tidying: ICPMS'
author: "James Vesto"
date: "12/16/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
```

```{r}
#Note! Before running this file you must run file 1 (01_ICPMS_data_tidying.Rmd)
ICPMS
```

1) Start by defining useful lists; such as what sites, metals, and sample IDs were analyzed.

```{r}
sample_sites <- unique(filter(ICPMS, site!= "QC", site!="MB", site!="")$site)
#excluding method blank and quality control from the list of sites
metals_analyzed <- unique(ICPMS$metal)

#Preview the lists to check for potential issues:
#sample_sites
#metals_analyzed
#unique_IDs #from file 1
```

2) Calibrating the ICPMS Data

```{r Calibration}

ICPMS_cal <- NULL

for (unique_metal in metals_analyzed){

  cal <- ICPMS %>% 
      filter(type == "Cal1" | type== "Cal2" | type == "Cal3") %>% 
      filter(metal == unique_metal) %>%
      select(conc = concentration, signal = CPS.corrected, RSD = RSD.corrected) #(Kg/mL)

  w <- 1/(cal$signal*cal$RSD)^2 
  model <- lm(cal$signal ~ cal$conc, weights= w)

  slope <- model$coefficients[2]
  intercept <- model$coefficients[1]
  slope_STD <- summary(model)$coefficients[2,2]
  intercept_STD <- summary(model)$coefficients[1,2]

  plot(cal$signal ~ cal$conc,
      xlab= paste("Concentration of ", unique_metal,"(ppb)"),
      ylab= "Counts per second")+ 
      abline(model, col="red")+ 
      title(paste("Calibration for", unique_metal)) 

  equation <- data.frame(metal= unique_metal, slope, slope_STD, intercept, intercept_STD)
  ICPMS_cal <- rbind(ICPMS_cal, equation)
  }

  ICPMS_cal 
  
  #Clearing the environment (optional, but helps to prevent accidentally using the wrong object!)
  remove(equation, cal, slope, slope_STD, intercept, intercept_STD, w, model, unique_metal)
```

2) Creating a function to analyze samples
```{r}
#inputs: unique_site (as a character, ex. "A")
#outputs: concentration vector (containing: Metal, site, )

sample_analysis <- function(unique_site){
  concentration_data <- NULL
  
  for (unique_metal in metals_analyzed){
    sample_data <- filter(ICPMS, metal == unique_metal, site == unique_site) 
    cal <- filter(ICPMS_cal, metal == unique_metal) 
  
    m <- cal$slope 
    b <- cal$intercept 
    y <- sample_data$CPS.corrected
  
    intercept_error <- cal$intercept_STD
    slope_error <- cal$slope_STD
  
    x <- (y-b)/m #The units are dependent on the calibration standards (Kg/mL)
    
    if(unique_site=="MB"){
      x <- mean(x)}
  
    concentrations <- data.frame(sample_ID = sample_data$sample.key,
                                metal= unique_metal,
                                site = unique_site,
                                concentration_diluted=x, 
                                intercept_error, 
                                slope_error)
    
    concentration_data <- rbind(concentration_data, concentrations)}
  return(concentration_data)}

```

3) Creating a run_sites function
```{r}
#inputs: a function
#outputs: a data frame with the function outputs from each site
run_sites <- function(Function){
  value <- NULL
  for(sites in sample_sites){
    site_value <- Function(sites)
    value <- rbind(site_value, value)}
  return(value)
  }
```

4) Running the calibration to extract the MB and sample concentrations post dilution
```{r}
MB <- sample_analysis("MB")
sample_conc <- run_sites(sample_analysis)
```

5) Method Blank Correction: subtracting concentrations of metals in the method blank from concentrations in the samples
```{r, warning=FALSE}
correction <- NULL

for (unique_metal in metals_analyzed){
  MB_metal <- filter(MB, metal==unique_metal)
  metal_sample_data <- mutate(sample_conc, concentration.dil.blanked = concentration_diluted-MB$concentration_diluted)
  correction <- rbind(correction, metal_sample_data)
  }
#This is producing a warning that longer object length is not a multiple of shorter object length!
#Need to investigate as to why this is happening and make sure it isnt affecting the results!
correction

remove(unique_metal, MB_metal, metal_sample_data)
```
