---
title: "03_ICPMS_Statistics_Visualization"
author: "James Vesto"
date: "7/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
```

##Note! Before running this file you must run files 1 and 2.
#This is the portion of the analysis where you get to think critically and creatively
#to perform visualizations, stastical analyses and tell a story about the data.

#It may be necessary to work back through the first to files to store more data based on the questions you want to ask, such as what had the largest contribution to the uncertainty in the results.

1) View Data
```{r}
sample_data
```
2) Average data for each site

```{r, message=FALSE}
#The message is: `summarise()` ungrouping output (override with `.groups` argument)
site_data <- NULL

for(sites in sample_sites){
 site_data <- filter(sample_data, site==sites)%>%
    group_by(metal)%>%
    summarise(site=sites,
              conc.blanked = mean(conc.blanked),
              conc.blanked.error = propmean(e_x = conc.blanked.error, x = conc.blanked),
              conc.unblanked = mean(conc.unblanked),
              conc.unblanked.error = propmean(conc.unblanked.error, conc.unblanked))%>%
   rbind(site_data)
}

site_data
```
