---
title: "03_ICPMS_Statistics_Visualization"
author: "James Vesto"
date: "7/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
```

##Note! Before running this file you must run files 1 and 2.
#This is the portion of the analysis where you get to think critically and creatively
#to perform visualizations, stastical analyses and tell a story about the data.

#It may be necessary to work back through the first to files to store more data based on the questions you want to ask, such as what had the largest contribution to the uncertainty in the results.

1) View Data
```{r}
sample_data
```
2) Average data for each site

```{r, message=FALSE}
#The message is: `summarise()` ungrouping output (override with `.groups` argument)
site_data <- NULL

for(sites in sample_sites){
 site_data <- filter(sample_data, site==sites)%>%
    group_by(metal)%>%
    summarise(site=sites,
              conc.blanked = mean(conc.blanked),
              conc.blanked.error = propmean(e_x = conc.blanked.error, x = conc.blanked),
              conc.unblanked = mean(conc.unblanked),
              conc.unblanked.error = propmean(conc.unblanked.error, conc.unblanked))%>%
   rbind(site_data)
}

site_data
```
3) Plot concentration data at each site
*Note: good practice for creating a visualization is to draw out approximately what you want it to look like on paper first. Then use the ggplot2 cheatsheet to help you create it in R!*

```{r}
#testing out a plot idea
site_data %>% filter(site=="A")%>%
  ggplot(aes(metal, conc.blanked))+ 
  geom_col()

#decided this wasnt the best decision, given that the concentration of lead extremely overshadows the other metals
```
```{r}
plot <- NULL
for(metals in metals_analyzed){
 data <- site_data %>%
   filter(metal==metals, site != "QC")
   
  plot <-  ggplot(data, aes(site, conc.blanked))+ 
    geom_col(fill="#abd7eb")+
    labs( title=paste("Concentration of", metals, "in Johnson Creek Watershed Sites"),
          y= paste("Concentration of", metals, "(ppm)"), #double check unit!!
          x="Site along Johnson Creek Watershed")+     
    geom_errorbar(ymin= data$conc.blanked - data$conc.blanked.error,
                  ymax= data$conc.blanked + data$conc.blanked.error)
    
 print(plot)
}

#This graph looks really neat but is perhaps a bit suspicious in the case of Pb208?
```
